---
title: "Peer Assignment_1 (RR)"
author: "Hitaxi"
date: "24/08/2020"
output: html_document
---

### Loading and Preprocessing Data

```{r, cache=TRUE}
activityData <- read.csv("activity.csv")
summary(activityData)

```
```{r}
names(activityData)
```
```{r}
head(activityData)
```

### Mean Total number of steps taken per day
#### Calculate total number number of steps taken per day and plot histogram
```{r, cache=TRUE}
stepsPerDay <- aggregate(steps ~ date, activityData, sum, na.rm=TRUE)
hist(stepsPerDay$steps)
```
#### Calculating the mean and meadian 
```{r, cache=TRUE}
meanStepsPerDay <- mean(stepsPerDay)
meanStepsPerDay
```

### Average activity daily pattern
```{r, cache=TRUE}
stepsPerInterval <- aggregate(steps ~ interval, data= activityData, mean, na.rm=TRUE )
plot(steps~interval, data=stepsPerInterval, type="l")
```
#### Calculating the 5-min interval maximum steps
```{r, cache=TRUE}
intervalWithMaxSteps <- stepsPerInterval[which.max(stepsPerInterval$steps),]$interval
intervalWithMaxSteps
```
### Imputing the missing values
```{r, cache=TRUE}
totalMissingValues <- sum(is.na(activityData$steps))
totalMissingValues
```
#### Filling the missing values with the mean of the data
```{r, cache=TRUE}
getMeanStepsPerInterval <- function(interval) {
  stepsPerInterval[stepsPerInterval$interval==interval,]$steps
}
activityData_1 <- activityData
for(i in 1:nrow(activityData)) {
  if(is.na(activityData_1[i,]$steps)) {
    activityData_1[i,]$steps <- getMeanStepsPerInterval(activityData_1[i,]$interval)
    }
  }
```
#### Histogram of total number of steps Per Day for new dataset
```{r, cache=TRUE}
totalStepsPerDayNoNA <- aggregate(steps ~date, data=activityData_1, sum)
hist(totalStepsPerDayNoNA$steps)
```
```{r, cache=TRUE}
meanSteps <- mean(totalStepsPerDayNoNA$steps)
medianSteps <- median(totalStepsPerDayNoNA$steps)
```
### Difference in activity patterns between and Weekdays and Weekends
```{r, cache=TRUE}
activityData_1$date <- as.Date(strptime(activityData_1$date, format = "%Y-%m-%d"))
activityData_1$day <- weekdays(activityData_1$date)
for (i in 1:nrow(activityData_1)) {
  if(activityData_1[i,]$day %in% c('Saturday','Sunday')) {
   activityData_1[i,]$day <- "weekend"
  }
  else {
    activityData_1[i,]$day <- "weekday"
  }
}
stepsbyDay <- aggregate(activityData_1$steps ~ activityData_1$interval + activityData_1$day, activityData_1,mean)
```
#### TO make plot containing the time-series
```{r, cache=TRUE}
names(stepsbyDay) <- c("interval","day","steps")
library(lattice)
xyplot(steps ~ interval |day, stepsbyDay, type="l", layout=c(1,2), xlab="Interval", ylab="Number of steps")
```













